Model Recognition as Planning

### 2019-02-27 doc/icaps-19/camera-ready/model-recognition.tex
ICAPS19: Minor corrections to section 4

Let be $\Psi$ the set of {\em predicates} that shape the propositional state variables $F$, and a list of {\em parameters} $pars(\xi)$. The set of elements that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the \strips\ action schema $\xi$ is given by FOL interpretations of $\Psi$ over the parameters $pars(\xi)$. We denote this set of FOL interpretations as ${\mathcal I}_{\Psi,\xi}$. For any of the six action schemata of Figure~\ref{fig:model-example}, the ${\mathcal I}_{\Psi,\xi}$ set contains the same ten elements, ${\mathcal I}_{\Psi,\xi}=${\small\tt\{xcoord($v_1$), xcoord($v_2$), ycoord($v_1$), ycoord($v_2$), q0(), q1(), next($v_1$,$v_1$), next($v_1$,$v_2$), next($v_2$,$v_1$), next($v_2$,$v_2$))\}}.

Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, the space of possible \strips\ schemata is constrained by a set ${\mathcal C}$ that includes:

@@

\begin{definition}[{\em Closest consistent models}] \label{consistent}
Given a model $\mathcal{M}$, the set $M^*$ of the {\bf closest consistent models} is the set of models $\mathcal{M}'$ that: (1) produce a trajectory $\tau(\pi,P)$ that reaches the goals in $P$ and is {\em consistent} with $\mathcal{O}(\tau)$ and (2) their {\em edit distance} to $\mathcal{M}$ is minimal;
  \[\underset{\forall \mathcal{M}' \rightarrow \mathcal{O}}{\arg\min}\ \delta(\mathcal{M},\mathcal{M}') \]
\end{definition}

@@

Provided that there is a single possible trajectory and a single possible model consistent with the input observation, i.e. $M^*=\{\mathcal{M}^*\}$ then, the probabilities of expression (2) are not added up and expression (1) simplifies to:
\begin{align}
argmax_{\mathcal{M}\in M} P(\mathcal{M^*}|\mathcal{M}) P(\mathcal{M}).
\end{align}
Note that the term $P(\tau|\mathcal{M^*})$ is taken out of the maximization because is independent of the input model $\mathcal{M}\in M$.

@@

Similar assumptions were taken in previous work for {\em goal recognition}~\cite{ramirez2012plan}. They allow us to assume that the sum in equation (2) is dominated by its largest term, so other terms in the sum are not added up. The largest term corresponds here to the shortest trajectory $\tau^*$ consistent with the observation (which now is assumed to be unique). Note that the more complete is the observation of the plan execution the more accurate is our estimate because it becomes more likely to assume that there is only one trajectory consistent with the input observation.

@@

We know from {\em pattern recognition}~\cite{devroye2013probabilistic} that, if string symbols are uniformly random and independent, the distance of a given string to a fixed string follows a {\em Binomial distribution}. Moreover, the probability that a particular string will be within a distance $D$ is the Cumulative Distribution Function (CDF) for the {\em Binomial distribution}~\cite{wilcox1981review}:
\begin{align}
CDF(D)=\sum_{d=0}^D{{N}\choose{d}}  p^d (1-p)^{N-d}
\end{align}
where $N$ is the length of the string, and $p<0.5$ since we consider that the cost of applying an {\em edit operation} is higher than do not applying it.

>>>>

Let $\Psi$ be the set of {\em predicates} that shape the propositional state variables $F$. The set of elements that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the \strips\ action schema $\xi$ is given by FOL interpretations of $\Psi$ over the parameters $pars(\xi)$. We denote this set of FOL interpretations as ${\mathcal I}_{\Psi,\xi}$. For any of the six action schemata of Figure~\ref{fig:model-example}, the ${\mathcal I}_{\Psi,\xi}$ set contains the same ten elements, ${\mathcal I}_{\Psi,\xi}=${\small\tt\{xcoord($v_1$), xcoord($v_2$), ycoord($v_1$), ycoord($v_2$), q0(), q1(), next($v_1$,$v_1$), next($v_1$,$v_2$), next($v_2$,$v_1$), next($v_2$,$v_2$))\}}.

Although any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, the space of possible \strips\ schemata is constrained by a set ${\mathcal C}$ that includes:

@@

\begin{definition}[{\em Closest consistent models}] \label{consistent}
Given a model $\mathcal{M}$, the set $M^*$ of the {\bf closest consistent models} is the set of models $\mathcal{M}'$ that: (1) produce a trajectory $\tau(\pi,P)$ that reaches the goals in $P$ and is {\em consistent} with $\mathcal{O}(\tau)$ and (2) their {\em edit distance} to $\mathcal{M}$ is minimal;
  \[M^*=\underset{\forall \mathcal{M}' \rightarrow \mathcal{O}}{\arg\min}\ \delta(\mathcal{M},\mathcal{M}') \]
\end{definition}

@@

Provided that there is a single possible trajectory and a single possible model consistent with the input observation, i.e. $M^*=\{\mathcal{M}^*\}$ then, the probabilities of expression (2) are not added up and expression (1) simplifies to:
\begin{align}
argmax_{\mathcal{M}\in M} P(\mathcal{M^*}|\mathcal{M}) P(\mathcal{M}).
\end{align}
Note that the term $P(\tau|\mathcal{M^*})$ is taken out of the maximization because it is independent of the input model $\mathcal{M}\in M$.

@@

Similar assumptions were taken in previous work for {\em goal recognition}~\cite{ramirez2012plan}. They allow us to assume that the sum in equation (2) is dominated by its largest term, so other terms in the sum are not added up. The largest term corresponds here to the shortest trajectory $\tau^*$ consistent with the observation (which now is assumed to be unique). Note that the more complete the observation of the plan execution is the more accurate our estimate is because it becomes more likely to assume that there is only one trajectory consistent with the input observation.

@@

We know from {\em pattern recognition}~\cite{devroye2013probabilistic} that, if string symbols are uniformly random and independent, the distance of a given string to a fixed string follows a {\em Binomial distribution}. Moreover, the probability that a particular string will be within a distance $D$ is the Cumulative Distribution Function (CDF) for the {\em Binomial distribution}~\cite{wilcox1981review}:
\begin{align}
CDF(D)=\sum_{d=0}^D{{N}\choose{d}}  p^d (1-p)^{N-d}
\end{align}
where $N$ is the length of the string, and $p<0.5$ since we consider that the cost of applying an {\em edit operation} is higher than not applying it.
